How are neurons organized into systems? 

\section{Neuroanatomy}

\subsection{Nervous System Structure}

\subsubsection{Structural Nervous System}

\begin{itemize}
    \item \textbf{Central Nervous System (CNS)}
    \begin{itemize}
        \item Brain
        \item Spinal Cord
    \end{itemize}
    \item \textbf{Peripheral Nervous System (PNS)}

\end{itemize}

\subsubsection{Functional Nervous System}

What are the `jobs' of the nervous system?

\begin{itemize}
    \item Somatic Nervous System
    \begin{itemize}
        \item Skeletal Muscles (Striated)
        \item Sensory information in
        \item Voluntary motion out
    \end{itemize}
    \item Autonomic Nervous System
    \begin{itemize}
        \item Uses smooth muscles
        \item Glands
        \item Sympathetic Nervous System
        \begin{itemize}
            \item Fight or Flight
            \item Heart rate, blood pressure, respiration, and alertness.
        \end{itemize}
        \item Parasympathetic Nervous System
        \begin{itemize}
            \item Rest and Digest
        \end{itemize}
        \item Enteric Nervous System
        \begin{itemize}
            \item A mesh-like system of neurons that governs the function of the gastrointestinal system.
            \item AKA: `Second Brain'
            \item GI problems are correlated with psychological disorders.
            \item The GI track houses a lot of our microbiota.
            \item Fecal Microbiota Transplant
            \begin{itemize}
                \item Rat studies showed that when a skinny rat has a fecal transplant from a fat rat, the skinny rat becomes fat. This works in reverse too.
                \item Therefore, the microbiota change the \textit{behavior} of the rat.
            \end{itemize}
            \item Elevated Plus Maze
            \begin{itemize}
                \item A test to measure anxiety in rats.
                \item The rats with the fecal transplant from the anxious rats were more anxious. 
                \item \textbf{This is huge!} This shows that the microbiota can change if a rat is anxious or not!
            \end{itemize}
        \end{itemize}
    \end{itemize}
\end{itemize}