\section{External Surface Structures}
\begin{itemize}
    \item \textbf{Ansate Sulcus / Cruciate Fissure} - Located on the dorsal surface, separating frontal and parietal lobes.
    \item \textbf{Medial Longitudinal Fissure} - Runs along the midline, separating the two cerebral hemispheres.
    \item \textbf{Cerebellum} - Located at the posterior base of the brain.
    \item \textbf{Coronal Sulcus / Superior Frontal Sulcus} - Found on the dorsal aspect, part of the frontal lobe.
    \item \textbf{Transverse Fissure} - Separates the cerebellum from the cerebrum.
    \item \textbf{Anterior / Rostral} - Toward the front of the brain.
    \item \textbf{Posterior / Caudal} - Toward the back of the brain.
    \item \textbf{Medial} - Toward the midline of the brain.
    \item \textbf{Lateral} - Away from the midline.
\end{itemize}

\section{Ventral Structures}
\begin{itemize}
    \item \textbf{Olfactory Bulb} - Located at the most anterior portion of the ventral side.
    \item \textbf{Cerebral Peduncle} - Found on the ventral midbrain, connects the cerebrum to the brainstem.
    \item \textbf{Pons} - Bulging structure on the brainstem, anterior to the medulla.
    \item \textbf{Pyramids on the Medulla} - Located on the ventral surface of the medulla oblongata.
    \item \textbf{Spinal Cord} - Extends from the medulla caudally.
    \item \textbf{Optic Chiasm} - X-shaped structure where optic nerves partially cross.
    \item \textbf{Median Eminence / Tuber Cinereum} - Located at the base of the hypothalamus.
    \item \textbf{Mammillary Body} - Small round structures part of the limbic system, posterior to the hypothalamus.
    \item \textbf{Periamygdaloid Cortex / Uncus} - Located in the medial temporal lobe, part of the piriform cortex.
    \item \textbf{Entorhinal Cortex} - Medial portion of the temporal lobe, key in memory processing.
    \item \textbf{Lateral Olfactory Tract} - Extends from the olfactory bulb along the ventral surface.
    \item \textbf{Rhinal Fissure} - Separates the neocortex from the piriform cortex.
    \item \textbf{Insula / Island of Reil} - Located deep within the lateral sulcus.
    \item \textbf{Sylvian Fissure} - Separates the temporal lobe from the frontal and parietal lobes.
\end{itemize}

\section{Midbrain and Hindbrain Structures}
\begin{itemize}
    \item \textbf{Superior Colliculus} - Dorsal midbrain, involved in visual processing.
    \item \textbf{Inferior Colliculus} - Below the superior colliculus, involved in auditory processing.
    \item \textbf{Pineal Body} - Small endocrine gland, located dorsally near the thalamus.
    \item \textbf{Hypothalamus} - Located below the thalamus, controls endocrine functions.
    \item \textbf{Corpus Callosum} - Large fiber bundle connecting the two cerebral hemispheres.
    \item \textbf{Cingulate Gyrus} - Located superior to the corpus callosum, involved in emotion regulation.
    \item \textbf{Cerebral Aqueduct} - Connects the third and fourth ventricles, running through the midbrain.
    \item \textbf{Tegmentum} - Ventral part of the midbrain, involved in motor control.
    \item \textbf{Massa Intermedia of the Thalamus} - Connects the two halves of the thalamus.
    \item \textbf{Fornix} - White matter tract connecting the hippocampus to the hypothalamus.
\end{itemize}

\section{Limbic System and Deep Brain Structures}
\begin{itemize}
    \item \textbf{Hippocampus} - Located in the medial temporal lobe, essential for memory formation.
    \item \textbf{Amygdala} - Anterior to the hippocampus, involved in emotion processing.
    \item \textbf{Fimbria} - White matter tract leading from the hippocampus.
    \item \textbf{Caudate Nucleus} - Part of the basal ganglia, involved in motor control.
    \item \textbf{Lateral Ventricle} - Space within the corpus callosum, part of the ventricular system.
    \item \textbf{Third Ventricle} - Space located between the two halves of the diencephalon.
\end{itemize}

\section{Cranial Nerves}
\begin{itemize}
    \item \textbf{Hypoglossal Nerve} - Located at the ventral medulla, controls tongue movement.
    \item \textbf{Glossopharyngeal / Vagus Nerve} - Found laterally on the medulla, involved in autonomic and sensory functions.
    \item \textbf{Auditory (Vestibulocochlear) / Facial Nerve} - Located near the pons, involved in hearing and facial expressions.
    \item \textbf{Abducens Nerve} - Controls lateral eye movement, located near the pons-medulla junction.
    \item \textbf{Trochlear Nerve} - Involved in eye movement, emerges dorsally from the brainstem.
    \item \textbf{Trigeminal Nerve} - Large nerve at the pons, responsible for facial sensation and chewing.
    \item \textbf{Oculomotor Nerve} - Controls most eye movements, located at the midbrain.
    \item \textbf{Optic Nerve} - Transmits visual information, connects to the optic chiasm.
\end{itemize}