\textbf{Chapter Outline:}
\begin{itemize}
    \item \hyperlink{hypothesis testing}{The Hypothesis-Testing Approach}
    \item The Case Study Method
    \item Statistical Analysis of Data
    \item Personality assessment
    \item Summary
\end{itemize}

\section{The Hypothesis-Testing Approach}\hypertarget{hypothesis testing}{}   
\begin{itemize}
    \item \textbf{Theory}: General statement about the relationship between constructs or events.
    \begin{itemize}
        \item Differ in the range of events or phenomena they cover.
    \end{itemize}
    \item Characteristics of a good theory:
    \begin{itemize}
        \item Parsimonious \textendash Explains the phenomenon in simple terms.
        \item Useful \textendash Generates testable hypotheses.
    \end{itemize}
    \item \textbf{Hypothesis}: Formal prediction about the relationship between or more variables that is logically derived from a theory.
    \item A theory is not accepted if empirical investigations consistently fail to confirm predictions.
\end{itemize}

\subsection{Types of Experimental Variables}

\begin{itemize}
    \item \textbf{Independent variable}: Manipulated by the experimenter.
    \item \textbf{Dependent variable}: Measured by the experimenter.
    \item \textbf{Non-Manipulated independent variable}: 
    \begin{itemize}
        \item Exists without the researcher's intervention.
        \item Investigator does not randomly assign participants to a conditions
        \item Research cannot assume the participants in the two groups are identical.
        \item Difficult to find cause-and-effect relationships.
    \end{itemize}
    \item \textbf{Manipulated independent variable}:
    \begin{itemize}
        \item Begins with numerous participants.
        \item Randomly assigns participants to experimental groups.
        \item Researcher can assume that all the differences will be evened out.
        \item Random assignment increases confidence in causation relationships.
    \end{itemize}
\end{itemize}

\subsubsection{Example}

\textbf{Theory}: Aggression is learned through observation. \\

\noindent Create a hypothesis based on the theory above. \\