\documentclass[11pt,letterpaper]{article}
\usepackage[
    ignoreheadfoot, % set margins without considering header and footer
    top=.6 in, % seperation between body and page edge from the top
    bottom=.6 in, % seperation between body and page edge from the bottom
    left=.75 in, % seperation between body and page edge from the left
    right=.75 in, % seperation between body and page edge from the right
    footskip=1.0 cm, % seperation between body and footer
    % showframe % for debugging 
]{geometry} % for adjusting page geometry
\usepackage{enumitem}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{parskip}
% \setlength{\parindent}{15pt} % Adjusts the amount of paragraph indentation
\setlength{\parskip}{10pt}    % Adds space between paragraphs

% Adjusting hyperlink color
\hypersetup{
    colorlinks=true,
    linkcolor=black,
    filecolor=black,      
    urlcolor=black,
    citecolor=black
}

% Custom formatting for header and footer
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}

\begin{document}

% Header with name and contact info
\begin{center}
    \LARGE \textbf{Paul Beggs} \\
    \vspace{0.2cm}
    \normalsize
    1600 Washington Ave \\
    Conway, AR 72032 \\
    \href{mailto:BeggsPA@hendrix.edu}{BeggsPA@hendrix.edu} \(\bullet\) (573) 587-9258 \\
    \url{https://www.linkedin.com/in/paul-beggs-0a8395224/}
\end{center}

\vspace{0.25cm}
\singlespacing




% Date and recipient information
\today

Melanie McCarty, LCSW \\
Conway Behavioral Health Hospital \\
2255 Sturgis Rd \\
Conway, AR 72034

% Opening
Dear Ms. McCarty,

% Introduction paragraph
% Explcitly state the position you are applying for. Explain why you are applying and then give a brief overview of what you will be discussing in the body of the letter.

I am writing to apply for the internship opportunity at Conway Behavioral Health Hospital. I have personal experience with rehabilitation which, in combination with my behavioral health and suicide intervention training, will allow me to effectively contribute to Conway Behavioral Health Hospital's mission of empowering patients to take active control of their recovery. Furthermore, Conway Behavioral Health Hospital's patient-centered care---rather than symptom-centered care---is, in my opinion, the most effective treatment for adverse behavioral health afflictions.

% Body paragraphs


Ever since I was introduced to therapeutic methods five years ago, I knew counseling was something I was going to do for the rest of my life. What started as a means for helping myself, soon turned into an overwhelming passion to further this practice so more people could be reached. I looked outward for volunteering institutions, so I could help where I could. Eventually, I found the Rape, Incest, and Abuse National Network hotline. I fostered my capacity to empathize, learn, and apply the skill set that I had been cultivating for years before I started training. Through these learned skills, I will help further the mission of Conway Behavioral Health Hospital.

In August, I underwent training to become a SOAR Mentor. This intensive program taught me how to identify signs of emotional distress, intervene appropriately, and collaborate with individuals to develop personalized safety plans (see r\'esum\'e). This experience not only reinforced my commitment to providing empathetic support but also deepened my understanding of effective intervention strategies. My ability to communicate compassionately, paired with my training in structured crisis response, equips me to contribute meaningfully to Conway Behavioral Health Hospital's patient-centered approach.



% Closing
% Wrap things up; thank them for their time and consideration, and give them some contact information in case they want to call you in for an interview.

In addition to my personal experience with rehabilitation and my training in behavioral health, I am confident that my passion for helping others make me an ideal candidate for the internship at Conway Behavioral Health Hospital. I am excited about the opportunity to contribute to your team and look forward to the possibility of discussing my application with you further. Thank you for considering my application. I hope to hear from you soon.


Sincerely, \\

\vspace*{1cm}

Paul Beggs


\end{document}
