\renewcommand{\theenumi}{\arabic{enumi}}
\renewcommand{\labelenumi}{\theenumi.}
\section{Introduction to the Power Set}

The \textit{Power Set} is an integral concept that handles the subsets of a set. That is, for any set \(A\), the power set \(\mathcal{P}(A)\) includes every possible subset of \(A\), ranging from the empty set to \(A\) itself. This section will delve into the formal structure of power sets, and provide a basis that we can build upon for later concepts.

\vspace{0.5cm}

\begin{axiom}
    {Power Set}Suppose $A$ is a set. Then, there exists a set $$\mathcal{P}(A)$$
    called the \textit{power set} of $A$ whose elements are exactly the subsets of $A$. Put formally, $$x\in \mathcal{P}(A) \iff x\subseteq A$$
\end{axiom}

\vspace{0.5cm}

\begin{note}
    \textbf{Note the following notation:}

\begin{enumerate}
        
    \item $x\in A$: This denotes that $x$ is an element of the set $A$. It means $x$ is one of the objects contained within the set $A$. For example, if $A=\{a,b,c\}$, then saying $a\in A$ is correct because $a$ is one of the elements in the set $A$.

    \item $x\subseteq A$: This denotes that $x$ is a subset of the set $A$. It means every element of $x$ is also an element of $A$, but $x$ itself does not need to be a singular 'element' in $A$; rather, it can be a collection (or set) of elements that are all found in $A$. For example, if $A=\{a,b,c\}$, then $\{a,b\} \subseteq A$ is correct because every element of the set $\{a,b\}$ is also an element of $A$.
\end{enumerate}

The key difference is that $x\in A$ refers to membership (i.e., $x$ is a single object within the collection $A$), whereas $x\subseteq A$ refers to inclusion (i.e., all elements of the collection $x$ are contained within the collection $A$).
\end{note}

\newpage

What we really need the power set for is to establish an ordering on a set. Up to now, membership was the only defining characteristic of sets, but now, we can specify a set's \textit{order} using the following notation for an \textit{ordered pair} $(x,y)$: $$\{\{x\},\{x,y\}\}$$

This leads us to the following Theorem:

\begin{ntheorem}
    {Existence pt. 3}Suppose $A$ and $B$ are sets with $a\in A$ and $b\in B$. Then the set $\{\{a\},\{a,b\}\}$ exists.
\end{ntheorem}

\tpf{
    Let $A$ and $B$ be sets such that $a\in A$ and $b\in B$. \\
    Then, $A\cup B$ exists, and $a,b\in A\cup B$ by the corollary to Axiom 1.1.1. 5. Consider $\mathcal{P}(A\cup B)$. By Problem 34, $\{a\}\in \mathcal{P}(P\cup B)$. And consider the set $a,b$, so $\{a,b\}\in \mathcal{P}(A\cup B)$ by Axiom 6. Consider $\mathcal{P}(\mathcal{P}(A\cup B))$ which exists by Axiom 6. Since $\{a\}, \{a,b\}\in \mathcal{P}(A\cup B)$, then $\{\{a\},\{a,b\}\} \in \mathcal{P}(\mathcal{P}(A\cup B))$.
}

\begin{ntheorem}
    {Set Pair Equality}Let $a,b,x,y$ are elements such that $\{\{a\},\{a,b\}\}=\{\{x\},\{x,y\}\}$. Then, $$a=x \text{ and } b=y$$ 
\end{ntheorem}

\tpf{
    Let $a,b,x,y$ be elements such that $\{\{a\},\{a,b\}\}=\{\{x\},\{x,y\}\}$. \\
    By Axiom 1, we must have $\{a\} = \{x\}$ or $\{a\} = \{x,y\}$. Consider the following cases: \\
    
    \noindent \textbf{Case 1:} \\
    \indent If $\{a\} = \{x\}$, then by Axiom 1 $a=x$. Also, in this case, $\{a,b\} = \{x,y\}$. And since $a=x$, then $b=y$ by Axiom 1. \\
    
    \noindent \textbf{Case 2:} \\
    \indent If $\{a\} = \{x,y\}$, then Axiom 1 tells us that either $a=x$ and $a=y$. Also in this case, the set containing $x$ must equal the set containing $a,b$ and thus, $x=a$ and $x=b$. \\
    
    Thus, $b=x=a=y$.
}

\begin{definition}
    {Ordered Pairs (formal definition)}Let $A$ and $B$ be the sets with $a\in A$ and $b\in B$. Then, the \textit{ordered pair} $(a,b)$ is the set $\{\{a\},\{a,b\}\}$.
\end{definition}

\newpage

\begin{ntheorem}
    {Product}Let $A$ and $B$ be sets. Then, $$\{(a,b)\colon a\in A \text{ and } b\in B\} \text{ exists.}$$ 
\end{ntheorem}

\tpf{
Let $A$ and $B$ be sets. Theorem 3.1.2 gives that each ordered pair exists, and the existence Axiom---simply stated as, ``There exists a set''---gives the existence of the set as a whole. 
}

\begin{definition}
    {Cartesian Product}The set in Theorem 4 is called the (Cartesian) product. We denote this as $A\times B$.
\end{definition}