\documentclass[11pt]{article}
\usepackage{draculatheme}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{geometry}
\linespread{1}

\begin{document}

\section{How to Write A Good Essay}

\subsection{Developing a Thesis}

Thesis statement needs to be narrow. Focus on an article by critiquing a prompt, question, etc. It needs to be \textit{argumentative}.

(Note: For a thesis idea, you can talk about mental disorders by using resources from the text.)

\subsection{Points to Focus On}

\begin{itemize}
    \item Need a hook.
    \begin{itemize}
        \item Like a poem, a work of art, something someone is going to recognize, an example of what your debate is going to focus on.
        \item Needs a good question: Uncontrollable wonder.
        \begin{itemize}
            \item ``Lead instead of telling.''
        \end{itemize}
        \item Provide answers for other questions that have already been asked.
        \begin{itemize}
            \item Summarize a literature review (optional).
        \end{itemize}
    \end{itemize}
    \item The return. 
    \begin{itemize}
        \item State thesis.
    \end{itemize}
    \item The body:
    \begin{itemize}
        \item Pick any of the following:
        \begin{itemize}
            \item Deepening the argument.
            \begin{itemize}
                \item Go into further detail about the thesis. Argue for novelty in your thesis.
            \end{itemize}
            \item Reckoning with the counterargument.
            \begin{itemize}
                \item A conflict / struggle. 
                \item Back and forth with the literature.
                \item Answer questions before they are asked. 
            \end{itemize}
            \item Disclosing.
            \begin{itemize}
                \item Truth, values, knowledge is equivalent to saying something correctly. This is wrong.
                \item We need to take secrets that are hidden from the normal view, and bring it to light.
                \item Created something new.
            \end{itemize}
        \end{itemize}
    \end{itemize}
    \item The return of the hook. 
    \item A moment of exhilaration.
    \begin{enumerate}
        \item Moments of ``Aha!'' to change the way someone perceives a concept.
        \item Involves:
        \begin{itemize}
            \item Feeling
            \item Change in emotion
            \item Articulation of what people understand
            \item Willing to change the readers' minds
        \end{itemize}
    \end{enumerate}
    \item End the essay with further work. That is, how can you show that the conversation can go further?
    \begin{itemize}
        \item Come up with objections.
    \end{itemize}
\end{itemize}



\end{document}